\documentclass[12pt]{exam}
% essential packages
\usepackage{fullpage} % margin formatting
\usepackage{enumitem} % configure enumerate and itemize
\usepackage{amsmath, amsfonts, amssymb, mathtools} % math symbols
\usepackage{xcolor, colortbl} % colors, including in tables
\usepackage{makecell} % thicker \Xhline in table
\usepackage{graphicx} % images, resizing
% sometimes needed packages
\usepackage{hyperref} % hyperlinks
% \hypersetup{colorlinks=true, urlcolor=blue}
% \usepackage{logicproof} % natural deduction
% \usepackage{tikz} % drawing graphs
% \usetikzlibrary{positioning}
% \usepackage{multicol}
% \usepackage{algpseudocode} % pseudocode
% paragraph formatting
\setlength{\parskip}{6pt}
\setlength{\parindent}{0cm}
% newline after Solution:
\renewcommand{\solutiontitle}{\noindent\textbf{Solution:}\par\noindent}
% less space before itemize/enumerate
\setlist{topsep=0pt}
% creates \filcl to grey out cells for groupwork grading
\newcommand{\filcl}{\cellcolor{gray!25}}
% creates \probnum to get the problem number
\newcounter{probnumcount}
\setcounter{probnumcount}{1}
\newcommand{\probnum}{\arabic{probnumcount}. \addtocounter{probnumcount}{1}}
% use roman numerals by default
\setlist[enumerate]{label={(\roman*)}}
% creates custom list environments for grading guidelines, question parts
\newlist{guidelines}{itemize}{1}
\setlist[guidelines]{label={}, left=0pt .. \parindent, nosep}
\newlist{gwguidelines}{enumerate}{1}
\setlist[gwguidelines]{label={(\roman*)}, nosep}
\newlist{qparts}{enumerate}{2}
\setlist[qparts]{label={(\alph*)}}
\newlist{qsubparts}{enumerate}{2}
\setlist[qsubparts]{label={(\roman*)}}
\newlist{stmts}{enumerate}{1}
\setlist[stmts]{label={(\roman*)}, nosep}
\newlist{pflist}{itemize}{4}
\setlist[pflist]{label={$\bullet$}, nosep}
\newlist{enumpflist}{enumerate}{4}
\setlist[enumpflist]{label={(\arabic*)}, nosep}
\printanswers
\newcommand{\prevhwnum}{0}
\newcommand{\hwnum}{1}
\begin{document}
%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%
\title{EECS 203: Discrete Mathematics\\
Winter 2024\\
Homework \hwnum{}}
\date{}
\author{}
\maketitle
\vspace{-50pt}
\begin{center}
\huge Due \textbf{Thursday, January 25th}, 10:00 pm\\
\Large No late homework accepted past midnight.\\
\vspace{10pt}
\large Number of Problems: $7+1$
\hspace{3cm}
Total Points: $100+20$
\end{center}
\vspace{25pt}
\begin{itemize}
\item \textbf{Match your pages!} Your submission time is when you upload the
file, so the time you take to match pages doesn't count against you.
\item Submit this assignment (and any regrade requests later) on Gradescope.
\item Justify your answers and show your work (unless a question says
otherwise).
\item By submitting this homework, you agree that you are in compliance with
the Engineering Honor Code and the Course Policies for 203, and that you are
submitting your own work.
\item Check the syllabus for full details.
\end{itemize}
\newpage
%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%
\section*{Individual Portion}
%1
\subsection*{\probnum Collaboration and Support [3 points]}
\begin{qparts}
\item Give the names and uniqnames of 2 of your EECS 203 classmates (these
could be members of your homework group or other classmates).
\item When you have questions about the course content, where can you ask them?
Where are \textit{you} most likely to ask questions?
\item Name one self-care action you plan to do this semester to maintain your
overall well-being.
\end{qparts}
\begin{solution}
	\begin{parts}
		\part Will Guo, guowill \\
			Hannah Hoang-Pham, hoangpha
		\part I would most likely ask them in person at the lectures, or through text.
		\part I want to take a lot of walks, I find them a nice source of stress relief.
	\end{parts}
\end{solution}
%2
\subsection*{\probnum Rock the Vote [12 points]}
Let $p$ and $q$ be the following propositions:
\begin{itemize}
\item $p$: The election has been decided.
\item $q$: The votes have been counted.
\end{itemize}
Express each of these propositions as an English statement:
\begin{qparts}
\item $\neg p\rightarrow\neg q$
\item $\neg q \lor (\neg p \land q)$
\end{qparts}
\begin{solution}
	\begin{qparts}
		\item ``If the election has not been decided, then the votes have not been counted."
		\item ``Either the votes have not been counted or both the election has not been decided and the votes have been counted." \\This can be simplified to ``Either the votes have not been counted or the election has not been decided."
	\end{qparts}
\end{solution}
%3
\subsection*{\probnum Negation Station [16 points]}
For each of the following propositions, give their negation in natural English.
Your answer should not contain the original proposition. That is, you shouldn't
negate it as ``It is not the case that ...'' or something similar.
\textbf{Note:} You do not need to show work besides your translation, but you may
earn partial credit if you do.
\begin{qparts}
\item $a$ is greater than 6 or at most 2.
\item $b$ is a perfect square, odd, and not divisible by 7.
\item $c$ is odd whenever it is prime and greater than 3.
\item If $d$ is divisible by 2, then it is even.
\end{qparts}
\begin{solution}
	\begin{qparts}
		\item $a$ is at most 6 and greater than 2.
		\item $b$ is not a perfect square, is not odd, or is divisible by 7.
		\item $c$ is not odd whenever it is prime and greater than 3.
		\item $d$ is divisible by 2 and not even.
	\end{qparts}
\end{solution}
%4
\subsection*{\probnum Lying and Politics [16 points]}
Imagine a world with two kinds of people: knights and knaves, where knights always
tell the truth and knaves always lie. There are three people A, B, and C, and one
of them is the city mayor.
\begin{itemize}
\item A says ``I am not the city mayor."
\item B says ``The city mayor is a knave."
\item C says ``All three of us are knaves."
\end{itemize}
Is the city mayor a knight or a knave? As part of your solution, determine
everything you can about whether A, B, and C are knights or knaves.
\begin{solution}
	C's statement cannot be true, since he calls himself a knave with his statement. Thus, he is a knave, and at most 2 of the people are knaves. Hence only one of A or B is a knave. \\
	If A is a knave, they are the city mayor. B is then telling the truth.
	If A is telling the truth, B cannot be a knave because B and C would both be knaves, and have the city mayor among them. This would make B's statement true, causing a contradiction. Thus B is a knight, making C the knavish city mayor. \\
	In conclusion, the city mayor is a knave, C is always a knave, and B is always a knight.
\end{solution}
%5
\subsection*{\probnum Is Equivalence Equivalent to Equality? [15 points]}
Show that $(b \rightarrow a) \land (c \rightarrow a)$ is logically equivalent to $\neg(b \lor c) \lor a$. If you use a truth table, be sure to state why the table
tells you what you claim. If you use logical equivalences, be sure to cite each
law you use.
\begin{solution}
	\begin{align*}
		(b \rightarrow a) \land (c \rightarrow a) \equiv& (\neg b \vee a) \wedge (\neg c \vee a) \tag*{implication breakout}\\
		\equiv& (\neg b \vee a \wedge \neg c) \vee (\neg b \vee a \wedge a) & \tag*{distributive property} \\
		\equiv& (\neg b \vee a \wedge \neg c) \vee (\neg b \vee a) & \tag*{$a \wedge a \equiv a$} \\
		\equiv& (\neg b \vee a) \vee (\neg b \vee a \wedge \neg c)  & \tag*{commutativity of or} \\
		\equiv& (a \vee \neg b) \vee (\neg b \vee a \wedge \neg c)  & \tag*{commutativity of or} \\
		\equiv& a \vee \neg b \vee a \wedge \neg c  & \tag*{$\neg b \vee \neg b \equiv \neg b$} \\
		\equiv& a \vee \neg b \vee a \wedge \neg c & \tag*{$\neg b \vee \neg b \equiv \neg b$} \\
		\equiv& a \vee \neg b \wedge \neg c  & \tag*{$a \vee a \equiv a$} \\
		\equiv& a \vee \neg (b \vee c)  & \tag*{DeMorgan's Law}
	\end{align*}
\end{solution}
%6
\subsection*{\probnum Deduce, Reuse, Recycle [20 points]}
Given that the following statements are \textbf{true}:
\[ (p \land r) \rightarrow q \hspace{0.5in} \: \neg q \hspace{0.5in} \: r \lor s \hspace{0.5in} \: q \lor r \]
For each of the propositions, $p,\ q,\ r,$ and $s$, state its truth value and
explain. If it cannot be found, briefly explain why.
\begin{solution}
	$q$: False, since $\neg q$ is true. \\
	$r$: True, since $q$ is false, so for $q \vee r$ to be true, $r$ must be true. \\
	$s$: Unknown, $r$ is true so $r \vee s$ will be true no matter what value $s$ has. \\
	$p$: False. Since $q$ is false, the condition of the first statement must be false. $r$ is true, so $p$ must make the $\wedge$ statement false.
\end{solution}
%7
\subsection*{\probnum Preposterous Propositions [18 points]}
Consider the following truth table, where $s$, $t$, and $w$ are unknown
propositions.
\begin{center}
\begin{tabular}{|c c c|| c | c | c |}
\hline
$p$ & $q$ & $r$ &
\hspace{0.8cm}$s$\hspace{0.8cm} & \hspace{0.8cm}$t$\hspace{0.8cm} & \hspace{0.8cm}
$w$\hspace*{0.8cm}\\\hline
T & T & T & F & T & F\\\hline
T & T & F & T & F & F\\\hline
T & F & T & F & T & T\\\hline
T & F & F & F & T & F\\\hline
F & T & T & F & T & T\\\hline
F & T & F & F & F & F\\\hline
F & F & T & F & T & T\\\hline
F & F & F & F & T & F\\\hline
\end{tabular}
\end{center}
Use the above truth table to answer the following questions. For each unknown
proposition, $s$, $t$, and $w$:
\begin{itemize}
\item Find an equivalent compound proposition using $p$, $q$, and/or $r$.
\item You may use \textbf{only} $\land$, $\lor$, $\neg$, and parentheses in
each of your answers.
\item You may use $p$, $q$, and $r$ \textbf{at most once} in each of your
answers.
\end{itemize}
\begin{solution}
	$s$: $p \wedge q \wedge \neg r$ \\
	$t$: $\neg q \wedge r$ \\
	$w$: $r \wedge \neg(p \wedge q)$
\end{solution}
\end{document}